---
title: "Reproducible Research: Peer Assessment 1"
output: 
html_document:
md_document:
keep_md: false   
---

```{r}
library("ggplot2")
library(dplyr)
library(xtable)
```
  
## Loading and preprocessing the data
# Loading the data
```{r Load data, echo = TRUE}
if(!file.exists("activity.csv")){
    activity <- read.csv("activity.csv", header = TRUE, sep = ",")
    summary(activity)
}else
{
    unzip("activity.zip")
    activity <- read.csv("activity.csv", header = TRUE, sep = ",")
    summary(activity)
}
```
  
## What is mean total number of steps taken per day?
  
```{r}
grpDate <- activity %>% 
           group_by(date) %>% 
           summarise(Total_steps=sum(steps,na.rm = TRUE))

Mean_Total_steps <- mean(grpDate$Total_steps, na.rm = TRUE)
Median_Total_steps <-median(grpDate$Total_steps, na.rm = TRUE)
# Make a plot 
hist(grpDate$Total_steps,xlab="Total steps", ylab="Count",main = "Total steps per date",breaks = 10)
abline(v=Mean_Total_steps,col = "blue",lwd=3)
abline(v=Median_Total_steps,col = "red",lwd=3)
legend(x="topright", legend=c("mean","median"), col=c("blue","red"), bty="n", lwd=3)
```
  
The mean and median of the total number of steps taken per day are:  
* Mean = `r Mean_Total_steps`  
* Median = `r Median_Total_steps`  
  
## What is the average daily activity pattern?
  
```{r}
grpInterval <- activity %>%
              group_by(interval) %>%
              summarise(Avrg=mean(steps,na.rm = TRUE))
MaxIntervalAvrg <- grpInterval[grpInterval$Avrg==max(grpInterval$Avrg,na.rm = TRUE),]
plot.new()
plot(grpInterval$interval,grpInterval$Avrg, type = 'l',
     xlab = 'Interval',ylab = 'Average by interval',main = 'Average by interval')
```
The 5-minute interval which contains the maximum number of steps is: `r MaxIntervalAvrg`  
  
## Imputing missing values  
  
Total of missing values.   
```{r}
Total_Missing_Value <- is.na(activity$steps)
table(Total_Missing_Value)
```
  
The missing values is raplace by the mean for that 5-minute interval. 
  
```{r}
activity_copy <- activity
for(ii in unique(activity_copy$interval))
{
activity_copy$steps[is.na(activity_copy$steps) & (activity_copy$interval == ii)] =  grpInterval$Avrg[grpInterval$interval==ii]
}
```

```{r}
grpDateN <- activity_copy %>% 
           group_by(date) %>% 
           summarise(Total_steps=sum(steps,na.rm = TRUE))

Mean_Total_stepsN <- mean(grpDateN$Total_steps, na.rm = TRUE)
Median_Total_stepsN <-median(grpDateN$Total_steps, na.rm = TRUE)
# Make a plot 
hist(grpDateN$Total_steps,xlab="Total steps", ylab="Count",main = "Total steps per date",breaks = 10)
abline(v=Mean_Total_steps,col = "blue",lwd=1)
abline(v=Median_Total_steps,col = "red",lwd=1)
abline(v=Mean_Total_stepsN,col = "blue",lty=5,lwd=3)
abline(v=Median_Total_stepsN,col = "red",lty=2,lwd=3)
legend(x="topright", legend=c("mean","median","new mean","new median"), col=c("blue","red","blue","red"), bty="n", lwd=c(1,1,3,3),lty = c(1,1,5,2))
```
# the new Mean and median steps for each date are:  
* Mean = `r Mean_Total_stepsN`  
* Median = `r Median_Total_stepsN` 


## Are there differences in activity patterns between weekdays and weekends?

```{r}
activity_copy$DayG <- factor(weekdays(as.Date(activity_copy$date)) %in% c("samedi","dimanche"), 
               labels=c("weekday","weekend"), ordered=FALSE)
grpInterval  <- activity_copy %>% 
                group_by(interval,DayG) %>% 
                summarise(mean=mean(steps))
p <- ggplot(grpInterval, aes(interval, mean))
p <- p + geom_line(stat = "identity") + facet_grid(DayG ~.)
p +labs(x="Interval", y="Average number of steps")
ggsave("instructions_fig/sample_panelplot.png")
```

